On the probability of finding nonphysical solutions through shadowing\tnoteref{1}

### 2020-12-13 main_els.tex
Add highlights and edit intro. Final JCP version.

%%Research highlights

\begin{highlights}
\item This paper highlights through rigorous examples that numerical solutions of chaotic dynamical systems are not guaranteed to produce the correct statistics or long-term behavior;
\item We show through examples that it is theoretically possible to significantly alter the statistics of a chaotic system, e.g. the earth's climate, by minor parameter perturbations.
\end{highlights}

\begin{keyword}
Chaotic dynamics \sep numerical simulations \sep scientific computing 
\sep shadowing sensitivity analysis %% keywords here, in the form: keyword \sep keyword

@@

Owing to the shadowing lemma, it is generally believed that numerical simulations of chaotic governing equations, which are small perturbations of true solutions, are shadowed by true solutions of the governing physics. However, if these shadowing solutions are nonphysical, numerical simulations do not produce the correct long-time averages. Furthermore, the computation of parametric derivatives or {\em linear response} of 
long-time averages by using the idea of shadowing \cite{qiqi-lss}, leads to incorrect values for those derivatives. Thus, the evidence of nonphysicality of shadowing presented in this paper urges us to reconsider the use of chaotic numerical simulations, and statistical sensitivity analysis on these simulations, which are critical for cross-disciplinary applications. These implications are particularly important for climate studies, where chaotic models are widely used to make long-term predictions. In particular, three broad implications of the evidence in this paper may be stated, the third of which harnesses the statistical inconsistency between slightly different models:
\begin{itemize}
    \item we cannot trust the long-time averages generated by numerical simulations of chaotic processes, even when we account for model uncertainties and statistical noise due to finite-time averaging;
    \item we cannot trust the sensitivities of long-time averages or statistical averages to small parameter changes computed by using shadowing-based methods;
	\item it is possible for two models separated by tiny changes in parameter values to have drastically different statistical behavior. This means that the Lorenzian butterfly (a small perturbation) can significantly alter the climate of Texas (statistical/long-term average), and not just its weather (short-term averages).
\end{itemize}
Our subject -- shadowing -- and other dynamical systems concepts used in this paper are introduced in detail, where they appear, so as to make our presentation accessible to computational scientists across disciplines. 
The paper is organized as follows. In section \ref{sec:shadowingIntro}, we introduce the concept of shadowing and its applications in the context of chaotic numerical simulations. In section \ref{sec:nonphysical}, we define the notion of physicality and give examples to illustrate the different ways in which to obtain nonphysical solutions. Section \ref{sec:shadowingnonphysical} is dedicated to several examples of perturbed tent maps that have nonphysical shadowing solutions. In the discussion in section \ref{sec:discussion}, we illustrate with an example that a small parameter perturbation can change statistics significantly. The supplementary material, section \ref{sec:supp}, contains the numerical procedure used to generate shadowing solutions.

@@

{\em Acknowledgments:} The authors are grateful to Prof. Tim Sauer and Prof. John Williams for comments on the paper.
\\
\textbf{Funding:} This work was supported by Air Force Office of Scientific Research Grant No. FA8650-19-C-2207.
\textbf{Conflict of Interest:} The authors declare that they have no conflict of interest.

>>>>

%%Research highlights
\begin{keyword}
Chaotic dynamics \sep numerical simulations \sep scientific computing 
\sep shadowing sensitivity analysis %% keywords here, in the form: keyword \sep keyword

@@

Owing to the shadowing lemma, it is generally believed that numerical simulations of chaotic governing equations, which are small perturbations of true solutions, are shadowed by true solutions of the governing physics. However, if these shadowing solutions are nonphysical, numerical simulations do not produce the correct long-time averages. Furthermore, the computation of parametric derivatives or {\em linear response} of 
long-time averages by using the idea of shadowing \cite{qiqi-lss}, leads to incorrect values of those derivatives. Thus, the evidence of nonphysicality of shadowing presented in this paper urges us to reconsider the use of chaotic numerical simulations, and statistical sensitivity analysis on these simulations, which are critical for cross-disciplinary applications. These implications are particularly important for climate studies, where chaotic models are widely used to make long-term predictions. In particular, three broad implications of the evidence in this paper may be stated, the third of which harnesses the statistical inconsistency between slightly different models:
\begin{itemize}
    \item we cannot trust the long-time averages generated by numerical simulations of chaotic processes, even when we account for model uncertainties and statistical noise due to finite-time averaging;
    \item we cannot trust the sensitivities of long-time averages or statistical averages to small parameter changes computed by using shadowing-based methods;
	\item it is possible for two models separated by minor changes in parameter values to have drastically different statistical behavior. This means that the Lorenzian butterfly (a small perturbation) can significantly alter the climate of Texas (statistical/long-term average), and not just its weather (short-term averages).
\end{itemize}
Our subject -- shadowing -- and other dynamical systems concepts used in this paper are introduced in detail, where they appear, so as to make our presentation accessible to computational scientists across disciplines. 
The paper is organized as follows. In section \ref{sec:shadowingIntro}, we introduce the concept of shadowing and its applications in the context of chaotic numerical simulations. In section \ref{sec:nonphysical}, we define the notion of physicality and give examples to illustrate the different ways in which to obtain nonphysical solutions. Section \ref{sec:shadowingnonphysical} is dedicated to several examples of perturbed tent maps that have nonphysical shadowing solutions. In the discussion in section \ref{sec:discussion}, we illustrate with an example that a small parameter perturbation can change statistics significantly. The supplementary material, section \ref{sec:supp}, contains the numerical procedure used to generate shadowing solutions.

@@

\textbf{Funding:} This work was supported by Air Force Office of Scientific Research Grant No. FA8650-19-C-2207.
\textbf{Conflict of Interest:} The authors declare that they have no conflict of interest.
